---
title: "Ass4 - ETC5523 - Relativity Humidity Melbourn"
output: html_notebook
---

Group 6: SIGMA
mbha0014	Mayunk Bharadwaj
pgre0007	Priscila Grecov

Presentation: TUE 4-6.00PM - order 6 - question Mitch

Our question:

4. How does relative humidity in Melbourne change throughout the **day** and **year**? **When** is it **most uncomfortable** to walk around the city (humidity **below 30%** or **above 60%**)?

```{r reading-datasets}
sensor_readings <- read.csv("Microclimate_Sensor_Readings.csv")
sensor_local <- read.csv("Microclimate_Sensor_Locations.csv")
```

We need just the sensors that measure the relative humidity that are the ones with sensor_id = 5b and sensor_id = 5b.EPA-1hr


```{r filtering-sensor}
unique(sensor_readings$sensor_id)

library(dplyr)
sensor_readingsRH <- sensor_readings %>% filter(sensor_id %in% c("5b", "5b.EPA-1h"))
```

```{r showing-newtable}
unique(sensor_readingsRH$sensor_id)
```

```{r}
sensor_readingsRH$date <- as.Date(sensor_readingsRH$local_time)
sensor_readingsRH$timeStamp <- as.POSIXct(sensor_readingsRH$local_time, format = "%Y/%m/%d %I:%M:%S %p")
sensor_readingsRH$hour <- as.numeric(format(sensor_readingsRH$timeStamp, "%H"))
sensor_readingsRH$month <- as.numeric(format(sensor_readingsRH$date, "%m"))
sensor_readingsRH$hourMin <- format(sensor_readingsRH$timeStamp, "%H:%M")
unique(sensor_readingsRH$hour)
unique(sensor_readingsRH$month)
unique(sensor_readingsRH$hourMin)

sensor_readingsRH$month <- factor(month.abb[sensor_readingsRH$month], levels = month.abb)
unique(sensor_readingsRH$month)

# library(chron)
# sensor_readingsRH$hourMin2 <- chron(times=sensor_readingsRH$hourMin, format = "h:m")
# sensor_readingsRH$hourMin2 <- format(strptime(sensor_readingsRH$hourMin2,"%H:%M:%S"),'%H:%M')
```

```{r}
apply(sensor_readingsRH, 2, function(x) sum(is.na(x)))
```

```{r}
sensor_readingsRH %>% 
  filter(is.na(value))
```

```{r}
unique(sensor_readingsRH$site_id)
```

```{r}
library(ggplot2)
library(ggridges)
library(tidyverse)
library(hrbrthemes)
library(viridis)

unique(sensor_readingsRH$date)
```

1ST) HOW RELATIVITY HUMIDITY CHANGE THROUGH THE YEAR - EXPLORING THE DATA BY MONTH OVER THE YEAR

```{r during-year}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(site_id,date) %>%
  summarise(mean_RH = mean(value), .groups = 'drop') %>%
  ggplot(aes(x=date, y=mean_RH, group=site_id, color=site_id)) +
  geom_line() +
  geom_line(aes(y=60), color="black", linetype = "dotted") +
  geom_line(aes(y=30), color="black", linetype = "dashed") +
  theme_ipsum() +
  xlab("") +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks  = "10 days") +
  theme(axis.text.x=element_text(angle=60, hjust=1, size = 8)) 
```
By the graph below, the different sensors show the same pattern over the year. Then, we can join all the sensors taken the average or median.

```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(date) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop') %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y=mean_RH), color="blue") +
  #geom_line(aes(y=median_RH), color="red") +
  geom_line(aes(y=70), color="red", linetype = "dashed") +
  geom_line(aes(y=30), color="red", linetype = "dashed") +
  theme_ipsum() +
  xlab("") +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks  = "10 days") +
  theme(axis.text.x=element_text(angle=60, hjust=1, size = 8)) 
```
```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(date) %>%
  summarise(mean_RH = median(value), .groups = 'drop') %>%
  ggplot(aes(x=date, y=mean_RH)) +
  geom_line() +
  geom_line(aes(y=70), color="red", linetype = "dashed") +
  geom_line(aes(y=30), color="red", linetype = "dashed") +
  theme_ipsum() +
  xlab("") +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks  = "10 days") +
  theme(axis.text.x=element_text(angle=60, hjust=1, size = 8)) 
```
```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(date) %>%
  summarise(median_RH = median(value), .groups = 'drop') %>%
  filter(median_RH <= 30)
```
Just three days during all the year presented daily average below 30%: 21/11/2019, 20/12/2019, 30/01/2020.

```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(site_id, month) %>%
  summarise(median_RH = median(value), .groups = 'drop') %>%
  ggplot(aes(x=month, y=median_RH, group=site_id, color=site_id)) +
  geom_line() +
  theme_ipsum() 
```


```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(month) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop') %>%
  ggplot(aes(x=month, group = 1)) +
  geom_line(aes(y=mean_RH), color="blue") +
  geom_line(aes(y=median_RH), color="red") +
  theme_ipsum() 
```
```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(month) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop') %>%
  filter(mean_RH >= 70 | median_RH >= 70)
```
The months of May to August are the worst moths where the RH averages are greater than 70%. November seems to be the month with the lowest RH average (but in this dataset we have just the 2nd half data for November).

```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(month) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop')
```
2ND) HOW RELATIVITY HUMIDITY CHANGE THROUGH THE DAY - EXPLORING THE DATA BY HOUR OVER THE DAY.

```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(site_id, hour) %>%
  summarise(mean_RH = mean(value), .groups = 'drop') %>%
  ggplot(aes(x=hour, y=mean_RH, group=site_id, color=site_id)) +
  geom_line() +
  theme_ipsum() +
  scale_x_continuous(breaks = seq(0,23,1))
```
```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(hour) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop') %>%
  ggplot(aes(x=hour)) +
  geom_line(aes(y=mean_RH), color="blue") +
  geom_line(aes(y=median_RH), color="red") +
  geom_line(aes(y=70), color="black", linetype = "dashed") +
  geom_line(aes(y=30), color="purple", linetype = "dashed") +
  theme_ipsum() +
  scale_x_continuous(breaks = seq(0,23,1))
```
```{r}
library(plotly)
facet1 <- sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(month, hour) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop') %>%
  ggplot(aes(x=hour)) +
  geom_line(aes(y=mean_RH), color="blue") +
  geom_line(aes(y=median_RH), color="red") +
  geom_line(aes(y=70), color="black", linetype = "dotted") +
  geom_line(aes(y=30), color="purple", linetype = "dotted") +
  facet_wrap(vars(month)) +
  scale_x_continuous(breaks = seq(0,23,1)) +
  theme(axis.text.x=element_text(angle=60, hjust=1, size = 7)) 

ggplotly(facet1)  %>%
        config(displayModeBar = F)

```

```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(month, hour) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop') %>%
  filter(mean_RH <= 30 | median_RH <= 30)
```
```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b.EPA-1h") %>%
  group_by(hour, month) %>%
  summarise(mean_RH = mean(value), median_RH = median(value), .groups = 'drop') %>%
  ggplot(aes(x=month,  group = 1)) +
  geom_line(aes(y=mean_RH), color="blue") +
  geom_line(aes(y=median_RH), color="red") +
  geom_line(aes(y=70), color="black", linetype = "dashed", size=0.5) +
  geom_line(aes(y=30), color="purple", linetype = "dashed", size=0.5) +
  theme(axis.text.x=element_text(angle=60, hjust=1, size = 7)) +
  facet_wrap(vars(hour)) 
```

```{r}
sensor_readingsRH %>%
  filter(sensor_id == "5b") %>%
  filter(!is.na(value)) %>%
  group_by(hourMin2) %>%
  summarise(mean_RH = mean(value), .groups = 'drop') %>%
  ggplot(aes(x=hourMin2)) +
  geom_line(aes(y=mean_RH), color="blue") +
  theme_ipsum() +
  scale_x_time()
```












